---
title: "Exploratory Data Analysis on Customer's Accounts"
subtitle: "Insights"  
author: 
  - "Panagiotis Stylianos"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["ninjutsu", "lucy", "my-style.css"]
    seal: false
    includes:
       after_body: "collapseoutput.js"
    nature:
      highlightStyle: github 
      highlightLanguage: r 
      highlightLines: true
      highlightSpans: false 
      countIncrementalSlides: false
      slideNumberFormat: '%current%/%total%'
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(leaflet)
library(lubridate)
library(visdat)
library(patchwork)
library(janitor)
# The dataset contains 3 months of transactions for 100 customers 
library(tidyverse)
xaringanExtra::use_panelset()

style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)

file <- readxl::read_xlsx("data/synthesised-transaction-dataset.xlsx", 
                          trim_ws = TRUE, 
                          col_types = c("text","numeric", "text", "text", "text", "text", "text", "text", "numeric",
                                        "text", "numeric", "date", "text", "numeric", "text", "text", "text", "numeric",
                                        "text", "text", "text", "text","text"))


```

class: title-slide, my-one-page-font 

# Exploring Customer's Accounts
 <br><br><br><br><br><br>
.left[
.green[Panagiotis Stylianos]
] 

.left[`r Sys.Date()`]

.footer[`r icon::fa('book')` Data@ANZ ]

---

## .blue[Data Concerns]

.panelset[

.panel[.panel-name[Summary]

> **.blue[14%]** of the dataset has missing values.

> **.blue[93%]** of the bpay_biller_code and merchant_code variables is missing.

> The rest of the important numerical variables are **.blue[100%]** present. 

> Location of client with account .blue["ACC-2901672282"] seems to be incorrect

> **.blue[8]** of the transactions exceed the amount of **.blue[5000]** under which most of the transactions fell.

> What stand out is that only .red[one] of these transactions is registered as **.blue[expenses]** with the rest being **.blue[salary payments]**.


]

.panel[.panel-name[Missing Data]


```{r,  fig.width=14, fig.height=6, fig.retina=3}
visdat::vis_miss(file) +
  theme(
    axis.text = element_text(size = 10)
  ) +
  labs(
    title = "% of Missing Data"
  ) +
  theme(
    plot.background = element_rect(fill = "cornsilk")
  )

```


]

.panel[.panel-name[Client's Location]

> Use the interactive map to get exta info per customer (in blue) and merchant (in orange).

```{r map, echo=FALSE, fig.width=14, fig.height=6, fig.retina=3}
file <- file %>% 
  separate(long_lat, into = c("long", "lat"), sep = "-") %>% 
  separate(merchant_long_lat, into = c("merch_long", "merch_lat"), sep = "-") %>% 
  mutate(long = as.numeric(long),
         lat = as.numeric(lat),
         merch_long = as.numeric(merch_long),
         merch_lat = as.numeric(merch_lat),
         date = lubridate::as_date(date),
         month = month(date),
         day = day(date),
         lat = lat * (-1),
         merch_lat = merch_lat * (-1),
         time = stringr::str_sub(extraction, start = 12, end = 19),
         hour = stringr::str_sub(time, start = 1, end = 2))



customer_map_data <- file %>% 
  select(account,long, lat) %>% 
  distinct()

merch_map_data <-  file %>% 
  select(merch_long, merch_lat, merchant_suburb) %>% 
  na.omit() %>% 
  distinct()
  
leaflet() %>% 
  addProviderTiles("Stamen.Terrain") %>% 
  addMarkers(lng = customer_map_data$long, lat = customer_map_data$lat, 
             label = customer_map_data$account) %>% 
  addCircleMarkers(lng = merch_map_data$merch_long, lat = merch_map_data$merch_lat, label = merch_map_data$merchant_suburb,
                   color = 'orangee', fill = T, fillColor = 'orange')

```

]

.panel[.panel-name[Outliers]

> Hover over the points to gain more insight about the transaction and the customer.

``` {r, fig.width=14, fig.height=5, fig.retina=3}
library(FNN)

amount_knn <- get.knn(file[, 'amount'])



file_amount <- file %>% 
  select(date, amount, balance, account, txn_description)


file_amount$knn_score <- rowMeans(amount_knn$nn.dist)

plotly::ggplotly(
file_amount %>% 
  ggplot(aes(date, amount, size = knn_score, color = txn_description, label = account)) +
  geom_point() +
  geom_hline(yintercept = 5000, lwd = 1.5, color = "red") +
  scale_size(guide = "none") +
  ggthemes::scale_color_colorblind(guide = "none") +
  theme_classic()
) 
```

]
]

---
## .blue[Average Transaction Amount per Type and Month]

.panelset[
.panel[.panel-name[Summary]

The next table summarizes the average transaction per type and month.

> The average transaction is consistent across all months for every type.

> Use the search option to identify specific transactions by their type or the month they occurred.

]

.panel[.panel-name[Plot]
```{r, echo=FALSE, cache = TRUE, fig.height = 2, fig.width=4, fig.retina=3}
DT::datatable(
file %>% 
  mutate(
    month = case_when(
      month == '8' ~ 'August',
      month == '9' ~ 'September',
      month == '10' ~ 'October'
    )
  ) %>% 
  group_by(txn_description, month) %>% 
  summarise(average_transaction = round(mean(amount, na.rm = TRUE),2)) %>% 
  arrange(txn_description) 
) %>% 
  DT::formatStyle(columns = colnames(.), fontSize = '50%')

```
]
]
---
## .blue[Further Insight]

.panelset[
.panel[.panel-name[Summary]


> Finally, I attempted to provide some additional insight based on the customer's data.

> All transaction types except salary payments show insignificant variacne with the exception of few outliers.

> All salary payments in the dataset occur between 10:00 and 17:00, thus allowing the bank to detect potential fraudulent transactions.

> The last slide classifies the customers based on gender to identify which customers spend a bigger proporiton of their salary, thus helping the bank to assess their credit risk in case they are interested in a loan.


]

.panel[.panel-name[Transactions Over Time]

```{r, fig.width=14, fig.height=6, fig.retina=3}
file %>% 
  mutate(
    month = case_when(
      month == '8' ~ 'August',
      month == '9' ~ 'September',
      month == '10' ~ 'October'
    ),
    month = factor(month, levels = c("August", "September", "October"))
  ) %>% 
  filter(amount > 100) %>% 
  ggplot(aes(month, amount, fill = txn_description)) +
  geom_boxplot() + 
  labs(
    title = "Distribution of transaction amount by type "
  ) +


file %>% 
  group_by(hour, txn_description) %>% 
  count() %>% 
  ggplot(aes(hour, n, fill = txn_description)) +
  geom_col() + 
  facet_wrap(~txn_description) +
  theme_classic() +
  ggthemes::scale_fill_economist() +
  labs(
    x = 'Hour',
    y = "Transaction Volume"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, size = 4)
  ) +
  labs(
    title = "Number of Transactions by hour and type"
  )

```

]
.panel[.panel-name[Client Profile]

```{r, fig.width=14, fig.height=6, fig.retina=3}
income <- file %>% 
  group_by(account, gender) %>% 
  select(txn_description, balance, amount, date, gender) %>% 
  filter(txn_description == "PAY/SALARY") %>% 
  summarise(total_payment = sum(amount)) %>% 
  arrange(desc(total_payment))

expenses <- file %>% 
  group_by(account) %>% 
  select(txn_description, balance, amount, date) %>% 
  filter(txn_description %in% c("POS", "SALES-POS", "PAYMENT")) %>% 
  summarise(total_expenses = sum(amount)) %>% 
  arrange(desc(total_expenses))

client_profile <- inner_join(income, expenses, by = "account") %>% 
  mutate(exp_perc_of_salary = total_expenses/total_payment * 100) %>% 
  arrange(desc(exp_perc_of_salary))


client_profile %>% 
  ggplot(aes(total_payment, y = ..density.., group = gender)) +
  geom_density(aes(fill = gender, color = gender )) +
  geom_histogram(alpha = 0.6, fill = "grey", binwidth = 1000) +
  theme_classic() +
  theme(
    legend.position = "none",
  
  ) +
  labs(
    title = "Salary distribution",
    x = "Salary",
    y = ""

  ) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  

client_profile %>% 
  ggplot(aes(total_expenses, y = ..density..)) +
  geom_density(aes(fill = gender, color = gender )) +
  geom_histogram(alpha = 0.6, fill = "grey", binwidth = 500) +
  theme_classic() +
  theme(
    legend.position = "bottom"
  ) +
  labs(
    title = "Expenses distribution",
    x = "Total Expenses",
    y = ""

  ) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +


client_profile %>% 
  ggplot(aes(exp_perc_of_salary, y = ..density..)) +
  geom_density(aes(fill = gender, color = gender )) +
  geom_histogram(alpha = 0.6, fill = "grey") +
  theme_classic() +
  theme(
    legend.position = "none"
  ) +
  labs(
    title = "Distribution of expenses as % of salary",
    x = "Expenses as percentage of Salary",
    y = ""
    
  ) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() 

```

]
]
